---
title: 'Data Wrangling: Zoning, Permits, ACS, Home Price'
author: "Cory Rand"
date: "January 10, 2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
#rm(list = ls())
library(data.table)
library(rgdal)
library(sp)
library(rgeos)
library(tidyverse)
library(magrittr)
library(here)
library(stringr)
library(tidycensus)
library(sf)
#get list of all files in directory
shp_files <- list.files(here("raw_data","zoning_data"))

read_shps <- function(x){
  #get directory of specific shapefile
  file <- paste0(here("raw_data", "zoning_data"), "/", x)
  #get layer name of shapfile in file directory
  layer <- list.files(file)[1] %>% substr(1, nchar(list.files(file)[1]) -4)
  #read in shapefile
  temp <- readOGR(file, layer = layer)
  #make all column names capitals
  names(temp) <- names(temp) %>% toupper()
  #convert ordinance dates from string to date format
  temp@data$EDIT_DATE <- temp@data$EDIT_DATE %>% as.Date()
  #assign shapefile name indicating its timeperiod
  name <- str_replace_all(layer, " ", "_")
  name <- substr(name, nchar(name) - 6, nchar(name))
  assign(name, temp, envir = .GlobalEnv)
}
  
#use custom function to read in all shapefiles
sapply(shp_files, read_shps) %>% invisible()

#fix a few things 
AUG2012 <- st_2012
AUG2012 <- spTransform(AUG2012, proj4string(JAN2019))
SEP2014 <- spTransform(SEP2014, proj4string(JAN2019))
rm(st_2012)

#read in ward data
#wards <- readOGR(here("raw_data", "Boundaries - Wards (2015-)"), layer = "geo_export_2308d057-9a78-43d2-94ac-9978dec82556")
#wards <- spTransform(wards, proj4string(SEP2016))

#read in neighborhood data
zillow_shp <- readOGR(here("raw_data", "ZillowNeighborhoods-IL"), layer = "ZillowNeighborhoods-IL")
zillow_shp <- zillow_shp[which(zillow_shp$City == "Chicago"),]
zillow_shp <- spTransform(zillow_shp, proj4string(JAN2019))
#read in landmark data
landmarks_shp <-  readOGR(here("raw_data", "Individual Landmarks"), layer = "geo_export_3c7a72b6-7cfa-4d1b-85bf-bc7d064e06e7")

landmarks_shp <- spTransform(landmarks_shp, proj4string(JAN2019))
```


```{r}
#get centroid of most recent shpfile
JAN2019@data$CENTROID <- gCentroid(JAN2019, byid = TRUE)

shp_files <- ls()[grep("20", ls())]

for(i in 1:length(shp_files)){
  temp <- get(shp_files[i])
  matches <- over(JAN2019$CENTROID, as(temp, "SpatialPolygons"))
  JAN2019@data[, shp_files[i]] <- temp$ZONE_CLASS[matches]
  JAN2019@data[, paste0(shp_files[i], "_type")] <- temp$ZONE_TYPE[matches]
}

#clean up environment
parcel_data <- JAN2019
rm(list = ls()[grep("20", ls())])

#match landmarks
landmarks_shp$landmark <- landmarks_shp$landmark %>% as.character()
landmarks_shp <- landmarks_shp[!is.na(landmarks_shp$landmark),]

landmarks_shp$CENTROID <- gCentroid(landmarks_shp, byid = TRUE)
landmarks_shp$PARCEL <- over(landmarks_shp$CENTROID, as(parcel_data, "SpatialPolygons"))

landmarks_shp <- landmarks_shp[!is.na(landmarks_shp$PARCEL),]
parcel_data$LANDMARKED <- NA
parcel_data$LANDMARKED[landmarks_shp$PARCEL] <- landmarks_shp$landmark

#match neighborhoods to parcels
neighborhood_matches <- over(parcel_data$CENTROID, as(zillow_shp, "SpatialPolygons"))
parcel_data$REGION <- zillow_shp$RegionID[neighborhood_matches]

#clean up environment
rm(landmarks_shp, neighborhood_matches)
```

```{r}
zoning_data <- parcel_data@data[,c(1, 8, 10, 21:55)]
type_cols <- names(zoning_data)[grep("type", names(zoning_data))]
zoning_data <- zoning_data[,!(names(zoning_data) %in% type_cols)]
#note, date is not always mid month, doing this for simplicity
names(zoning_data)[grep("20", names(zoning_data))] <- paste0('15', names(zoning_data)[grep("20", names(zoning_data))])

#put data in long format
zoning_data <- gather(zoning_data, key = "Date", value = "Zoning", -c("OBJECTID", "REGION", "LANDMARKED", "CREATE_DAT", "EDIT_DATE"))

#convert to dates
zoning_data <- mutate(zoning_data, Date=as.Date(Date, "%d%b%Y"),
                                   LANDMARKED = as.Date(LANDMARKED, format = "%m/%d/%Y"),
                                   REGION = as.character(REGION), 
                                   CREATE_DAT = as.Date(CREATE_DAT))

#create floor area ratio data
Zoning <- c("RS-1", "RS-2", "RS-3", "RT-3.5", "RT-4", "RT-4A", "RM-4.5", "RM-5", "RM-5.5", "RM-6", "RM-6.5", "DR-10", "DR-3", "DR-5", "DR-7")
#note: RT4-A is 1.2-1.5, but here counting as 1.5
FAR <- c(0.5, 0.65, 0.9, 1.05, 1.2, 1.5, 1.7, 2, 2.5, 4.4, 6.6, 10, 3, 5, 7)
far_levels <- data.frame(Zoning, FAR, stringsAsFactors = F)

#merge floor area ratio data
zoning_data <- left_join(zoning_data, far_levels, by = "Zoning") %>%
                  subset(select = c(OBJECTID, REGION, Date, Zoning, FAR, LANDMARKED, CREATE_DAT, EDIT_DATE))

zoning_data <- zoning_data[order(zoning_data$LANDMARKED, zoning_data$OBJECTID),]

#lag the timeseries data
zoning_data_lag <- group_by(zoning_data, OBJECTID) %>%
                mutate(LAG_FAR = dplyr::lag(FAR, n = 1, default = NA_character_, order_by = Date))

#get landmark dummies
landmarked_parcels <- filter(zoning_data_lag, LANDMARKED > 0)
landmarked_parcels <- landmarked_parcels[!duplicated(landmarked_parcels[,1:2]),]
landmarked_parcels %<>% mutate(Date = LANDMARKED, LANDMARKED = 1)

#merge landmark dummy obs to parcel dataset
zoning_data_lag$LANDMARKED <- 0 %>% as.numeric()
zoning_data_lag %<>% bind_rows(landmarked_parcels)

#create upzone and downzone dummies
zoning_data_lag$UPZONE <- ifelse(zoning_data_lag$FAR > zoning_data_lag$LAG_FAR, 1, 0)


zoning_data_lag$DOWNZONE <- ifelse(zoning_data_lag$FAR < zoning_data_lag$LAG_FAR, 1, 0)

write_csv(zoning_data_lag, here("clean_data", "zoning.csv"))



########## do similar process but for zone type ###############
zoning_data <- parcel_data@data[,c(1, 8, 10, 21:55)]
type_cols <- names(zoning_data)[grep("type", names(zoning_data))]
zoning_data <- zoning_data[,c("OBJECTID", "CREATE_DAT", "EDIT_DATE", "REGION", type_cols)]
#note, date is not always mid month, doing this for simplicity
names(zoning_data)[grep("20", names(zoning_data))] <- gsub("_type", "", names(zoning_data)[grep("20", names(zoning_data))])
names(zoning_data)[grep("20", names(zoning_data))] <- paste0('15', names(zoning_data)[grep("20", names(zoning_data))])

#put data in long format
zoning_data <- gather(zoning_data, key = "Date", value = "Zoning_class", -c("OBJECTID", "REGION", "CREATE_DAT", "EDIT_DATE"))

#convert to dates
zoning_data <- mutate(zoning_data, Date=as.Date(Date, "%d%b%Y"),
                                   REGION = as.character(REGION), 
                                   CREATE_DAT = as.Date(CREATE_DAT),
                                   Zoning_class = as.numeric(Zoning_class))

#add in zone type codes
Zoning_type <- c("Business", "Commerical", "Manufacturing", "Residential", "Planned_Development", "Planned_Man_Districts", "Downtown_Mixed", "Downtown_Core", "Downtown_Res", "Downtown_Service", "Transportation", "Parks_Open_Space", NA)

#note: RT4-A is 1.2-1.5, but here counting as 1.5
codes <- c(1:12, 0)
zoning_type <- data.frame(Zoning_type, codes, stringsAsFactors = F)

#merge codes with zoning data
zoning_data <- left_join(zoning_data, zoning_type, by = c("Zoning_class" = "codes")) %>%
                  subset(!is.na(Zoning_type))

write_csv(zoning_data, here("clean_data", "zoning_type.csv"))
```


NOTES:
can't use AUG2012, no zone_class. Still gets read in by is nulled out when overlaid
There duplicates of Apr 2017, so I deleted on from raw data files
June 2016 actually Aug 2015, so I changed file name
June 2015 is also actually Aug 2015, so I deleted it.
June 2018 is July 2018, so I renamed it
2017-2 inner files were labeled Dec2016, but they are 2017-2 so I renamed them


Zone type is broad categories of zoning:
1 = B, Business
2 = C, Commercial
3 = M, Manufacturing
4 = R, Residential
5 = PD, Planned Development
6 = PMD, Planned Manufacturing Districts
7 = DX, Downtown Mixed Use
8 = DC, Downtown Core 
9 = DR, Downtown Residential
10 = DS, Downtown Service
11 = T, Transportation
12 = POS, Parks and Open Space
0 = NA

##############################
## ACS DATA
##############################
```{r Brown_ACS}
### 2010 DATA
brown_10 <- read_csv(here("raw_data", "acs_data", "LTDB_Std_All_fullcount", "LTDB_Std_2010_fullcount.csv"), col_names = T) %>%
            filter(state == "IL" & county == "Cook County")
names(brown_10) <- gsub("10", "", names(brown_10)) %>% tolower()
brown_10 %<>% select("tractid", "pop", "nhwht", "nhblk", "asian", "hisp", "hu", "vac", "ohu", "own", "rent" )
brown_10$YEAR <- "2010"
names(brown_10)[1] <- "trtid10"
brown_10$trtid10 %<>% as.character()

### 2000 DATA
brown_00 <- read_csv(here("raw_data", "acs_data", "LTDB_Std_All_fullcount", "LTDB_Std_2000_fullcount.csv"), col_names = T) %>%
            filter(state == "IL" & county == "Cook County")
names(brown_00) <- gsub("00", "", names(brown_00)) %>% tolower()
brown_00 %<>% select("trtid10", "pop", "nhwht", "nhblk", "asian", "hisp", "hu", "vac", "ohu", "own", "rent" )
brown_00$YEAR <- "2000"
brown_00$trtid10 %<>% as.character()

### 1990 DATA
brown_90 <- read_csv(here("raw_data", "acs_data", "LTDB_Std_All_fullcount", "LTDB_Std_1990_fullcount.csv"), col_names = T) %>%
            filter(state == "IL" & county == "Cook County")
names(brown_90) <- gsub("90", "", names(brown_90)) %>% tolower()
brown_90 %<>% select("trtid10", "pop", "nhwht", "nhblk", "asian", "hisp", "hu", "vac", "ohu", "own", "rent" )
brown_90$YEAR <- "1990"
brown_90$trtid10 %<>% as.character()

### 1980 DATA
brown_80 <- read_csv(here("raw_data", "acs_data", "LTDB_Std_All_fullcount", "LTDB_Std_1980_fullcount.csv"), col_names = T) %>%
            filter(state == "IL" & county == "Cook County")
names(brown_80) <- gsub("80", "", names(brown_80)) %>% tolower()
brown_80 %<>% select("trtid10", "pop", "nhwht", "nhblk", "asian", "hisp", "hu", "vac", "ohu", "own", "rent" )
brown_80$YEAR <- "1980"
brown_80$trtid10 %<>% as.character()

### 1970 DATA
brown_70 <- read_csv(here("raw_data", "acs_data", "LTDB_Std_All_fullcount", "LTDB_Std_1970_fullcount.csv"), col_names = T) %>%
            filter(state == "IL" & county == "Cook County")
names(brown_70) <- gsub("70", "", names(brown_70)) %>% tolower()
brown_70 %<>% select("trtid10", "pop", "white", "black", "asian", "hu", "vac", "ohu", "own", "rent" )
brown_70$YEAR <- "1970"
names(brown_70)[3:4] <- c("nhwht", "nhblk")

brown_acs <- bind_rows(brown_70, brown_80, brown_90, brown_00, brown_10)


zillow_shp <- readOGR(here("raw_data", "ZillowNeighborhoods-IL"), layer = "ZillowNeighborhoods-IL")

tracts_2010 <- readOGR(here("raw_data", "Boundaries - Census Tracts - 2010"), layer = "geo_export_52dfa5f1-1eaf-49e1-9862-1998098d508c") %>% spTransform(proj4string(zillow_shp))
tracts_2010$CENTROID <- gCentroid(tracts_2010, byid = T)
tract_matches <- over(tracts_2010$CENTROID, as(zillow_shp, "SpatialPolygons"))
tracts_2010$REGION <- zillow_shp$RegionID[tract_matches] %>% as.character()
tract_2010_data <- tracts_2010@data[,c(4, 11)]

brown_acs %<>% inner_join(tract_2010_data, by = c("trtid10" = "geoid10"))

brown_acs %<>% mutate(pwhite = nhwht/pop,
                     pblack = nhblk/pop,
                     pasian = asian/pop,
                     phisp = hisp/pop,
                     pohu = ohu/hu,
                     pown = own/hu,
                     pop_d_h = pop/hu,
                     h_d_pop = hu/pop)
                     
```


```{r}
sample_1970 <- read_csv(here("raw_data", "acs_data", "LTDB_Std_All_Sample", "ltdb_std_1970_sample.csv"), col_names = T, col_types = cols(.default = "c")) %>%
filter(state == "IL" & county == "Cook County") %>%
  select("TRTID10", "COL70", "UNEMP70", "FAMILY70", "HINC70", "MHMVAL70", "H30OLD70", "MRENT70")
names(sample_1970) <- gsub("70", "", names(sample_1970)) %>% tolower()
sample_1970$YEAR <- 1970


sample_1980 <- read_csv(here("raw_data", "acs_data", "LTDB_Std_All_Sample", "ltdb_std_1980_sample.csv"), col_names = T, col_types = cols(.default = "c")) %>% 
  filter(state == "IL" & county == "Cook County")
names(sample_1980) %<>% toupper()
sample_1980 %<>% select("TRTID10", "COL80", "UNEMP80", "FAMILY80", "HINC80", "H30OLD80", "HINCW80", "HINCB80", "HINCH80", "MULTI80")
names(sample_1980) <- gsub("80", "", names(sample_1980)) %>% tolower()
sample_1980$YEAR <- 1980

sample_1990 <- read_csv(here("raw_data", "acs_data", "LTDB_Std_All_Sample", "ltdb_std_1990_sample.csv"), col_names = T, col_types = cols(.default = "c")) %>%
  filter(state == "IL" & county == "Cook County") %>%
  select("TRTID10", "COL90", "UNEMP90", "FAMILY90", "HINC90", "H30OLD90", "HINCW90", "HINCB90", "HINCH90", "MULTI90")
names(sample_1990) <- gsub("90", "", names(sample_1990)) %>% tolower()
sample_1990$YEAR <- 1990


sample_2000 <- read_csv(here("raw_data", "acs_data", "LTDB_Std_All_Sample", "ltdb_std_2000_sample.csv"), col_names = T, col_types = cols(.default = "c")) %>%
  filter(state == "IL" & county == "Cook County") %>%
  select("TRTID10", "COL00", "UNEMP00", "HINC00", "H30OLD00", "HINCW00", "HINCB00", "HINCH00", "MULTI00")
names(sample_2000) <- gsub("00", "", names(sample_2000)) %>% tolower()
sample_2000$YEAR <- 2000


sample_2010 <- read_csv(here("raw_data", "acs_data", "LTDB_Std_All_Sample", "ltdb_std_2010_sample.csv"), col_names = T, col_types = cols(.default = "c")) %>% 
  filter(statea == "17" & countya == "031")
names(sample_2010) %<>% toupper()
sample_2010 %<>% mutate(TRACTID = paste0(STATEA, COUNTYA, TRACTA))
sample_2010 %<>% select("TRACTID", "COL12", "UNEMP12", "HINC12", "FAMILY12", "H30OLD12", "HINCW12", "HINCB12", "HINCH12", "MULTI12")
names(sample_2010)[1] <- "TRTID10"
names(sample_2010) <- gsub("12", "", names(sample_2010)) %>% tolower()
sample_2010$YEAR <- 2010

sample_full <- bind_rows(sample_1970, sample_1980, sample_1990, sample_2000, sample_2010) %>% 
                mutate(trtid10 = as.character(trtid10),
                       YEAR = as.character(YEAR),
                       col = as.numeric(col),
                       unemp = as.numeric(unemp),
                       family = as.numeric(family),
                       hinc = as.numeric(hinc),
                       mhmval = as.numeric(mhmval),
                       h30old = as.numeric(h30old),
                       mrent = as.numeric(mrent),
                       hincw = as.numeric(hincw),
                       hincb = as.numeric(hincb),
                       hinch = as.numeric(hinch),
                       multi = as.numeric(multi))

acs_full <- inner_join(brown_acs, sample_full, by = c("trtid10", "YEAR"))
write_csv(acs_full, here("clean_data", "acs_full.csv"))
```



```{r}
#v16 <- load_variables(year = 2017, dataset = "acs5")

acs_wide_2013_2017 <-  get_acs(geography = "tract", year = 2017, 
                 variables = c("B01003_001", "B25008_002", "B25008_003",
                               "B02001_002", "B02001_003", "B02001_004",
                               "B02001_005", "B03002_012", "B17026_001",
                               "B19025_001", "B25001_001", "B25003_001", "B25003_002", "B25003_003"), 
                 state = "IL", county = "Cook", geometry = F, output = "wide")

variable <-c("B01003_001E", "B25008_002E", "B25008_003E","B02001_002E", "B02001_003E", "B02001_004E","B02001_005E", "B03002_012E", "B17026_001E","B19025_001E", "B25001_001E", "B25003_001E", "B25003_002E", "B25003_003E",
             "B01003_001M", "B25008_002M", "B25008_003M","B02001_002M", "B02001_003M", "B02001_004M","B02001_005M", "B03002_012M", "B17026_001M","B19025_001M", "B25001_001M", "B25003_001M", "B25003_002M", "B25003_003M")
names <- c("pop_e", "owner_pop_e", "renter_pop_e", "white_pop_e", "black_pop_e", "indian_pop_e", "asian_pop_e", "hispanic_pop_e", "inc_pov_ratio_e", "agg_income_e", "housing_e", "occupied_e", "owned_e", "rented_e",
           "pop_m", "owner_pop_m", "renter_pop_m", "white_pop_m", "black_pop_m", "indian_pop_m", "asian_pop_m", "hispanic_pop_m", "inc_pov_ratio_m", "agg_income_m", "housing_m", "occupied_m", "owned_m", "rented_m")
var_names <- data.frame(variable, names, stringsAsFactors = F)

variable <- names(acs_wide_2013_2017) %>% data.frame(stringsAsFactors = F)
names(variable) <- "variable"
renames <- left_join(variable, var_names, by = "variable")

names(acs_wide_2013_2017)[3:30] <- renames$names[3:30]

acs_wide_2013_2017 <- inner_join(acs_wide_2013_2017, tract_2010_data, by = c("GEOID"="geoid10"))
write_csv(acs_wide_2013_2017, here("clean_data", "acs_2013_2017.csv"))
```

##############################
## Building Permits
##############################
```{r}
neighborhoods <- st_read("raw_data/ZillowNeighborhoods-IL/ZillowNeighborhoods-IL.shp") %>%
  st_set_crs(4326)

permits_raw <- read_csv(here("raw_data", "Building_Permits.csv"), col_names = T)
permits_narrow <- permits_raw[, c(1:13, 129:131)]
permits_narrow %<>% mutate(ISSUE_DATE = as.Date(ISSUE_DATE, format = "%m/%d/%Y"),
                           LOCATION = gsub(c("\\(|\\)"), "", LOCATION)) %>% 
                  filter(!is.na(LATITUDE) & !is.na(LONGITUDE)) %>%
                  st_as_sf(coords = c("LONGITUDE", "LATITUDE"), crs = 4326)

permits_narrow <- st_join(permits_narrow, neighborhoods %>% select(RegionID), join = st_within)

permits_filt <- filter(permits_narrow, PERMIT_TYPE == "PERMIT - NEW CONSTRUCTION" | PERMIT_TYPE == "PERMIT - WRECKING/DEMOLITION")

permits_filt$geometry <- NULL

write_csv(permits_filt, here("clean_data", "permits.csv"))


#categorize permits based on work descriptions
duplicate_words <- c("TEMP", "TEMPORARY", "CRANE", "REVISION", "EXPIRES", "HOIST", "FOUNDATION ONLY") %>% paste( collapse = "|")

residential_words <- c("DWELLING", "UNIT", "DU", "D.U.", "RESIDENCE", "RESIDENTIAL", "TOWNHOUSE", "TOWNHOME", "SFR",  "FAMILY", "MIXED", "MULTI", "HOUSE", "HOUSING", "RESIDENCY", "RESIDANCE", "APARTMENT", "FLAT", "FAMILEY") %>% paste( collapse = "|")

build_words <- c("ERECT", "EREC", "NEW", "CONSTRUCT") %>% paste( collapse = "|")
alter_words <-c("ALTER", "EXISTING", "REPAIR") %>% paste( collapse = "|")

single_family_words <- c("SINGLE FAMILY", "SFR", "RANCH", "FARM", "RESIDENCE") %>% paste( collapse = "|")
multi_family_words <- c("MIXED", "MULTI", "UNIT", "DU", "APARTMENT", "FLAT") %>% paste( collapse = "|")
townhome_words <- c("TOWNHOME", "TOWNHOUSE", "TOWNHOUSES") %>% paste( collapse = "|")

permits_filt %<>% mutate(YEAR = year(ISSUE_DATE),
                    REGION = as.character(RegionID),
                    WORK_DESCRIPTION =toupper(WORK_DESCRIPTION))
#get rid of most accessory permits
permits_filt <- permits_filt[!str_detect(permits$WORK_DESCRIPTION, regex(duplicate_words)),] %>%
      distinct(LOCATION, YEAR, PERMIT_TYPE, WORK_DESCRIPTION, .keep_all = T) 

permits_filt %<>% 
        mutate(
               WORK_DESCRIPTION = gsub("-", " ", WORK_DESCRIPTION),
               WORK_DESCRIPTION = gsub("[.]", "", WORK_DESCRIPTION),
               WORK_DESCRIPTION = gsub("ONE", "1", WORK_DESCRIPTION),
               WORK_DESCRIPTION = gsub("TWO", "2", WORK_DESCRIPTION),
               WORK_DESCRIPTION = gsub("THREE", "3", WORK_DESCRIPTION),
               WORK_DESCRIPTION = gsub("FOUR", "4", WORK_DESCRIPTION),
               WORK_DESCRIPTION = gsub("FIVE", "5", WORK_DESCRIPTION),
               WORK_DESCRIPTION = gsub("SIX", "6", WORK_DESCRIPTION),
               WORK_DESCRIPTION = gsub("SEVEN", "7", WORK_DESCRIPTION),
               WORK_DESCRIPTION = gsub("EIGHT", "8", WORK_DESCRIPTION),
               WORK_DESCRIPTION = gsub("NINE", "9", WORK_DESCRIPTION),
                      #1. Includes Residential (Dummy)
               RESIDENTIAL = str_detect(permits_filt$WORK_DESCRIPTION, regex(residential_words)),                             #1. Action
              BUILD = str_detect(permits_filt$WORK_DESCRIPTION, regex(build_words)) & 
                       PERMIT_TYPE =="PERMIT - NEW CONSTRUCTION",
              WRECK = PERMIT_TYPE =="PERMIT - WRECKING/DEMOLITION",
              DECONVERT = str_detect(permits_filt$WORK_DESCRIPTION, "DECONVERT"),
              ALTER = str_detect(permits_filt$WORK_DESCRIPTION, regex(alter_words)) &
                       BUILD != TRUE & WRECK != TRUE & DECONVERT != TRUE)

permits_filt %<>% mutate(
                #2. Identify Single vs Multi Family vs Townhome vs COACH
                          SFR = str_detect(permits_filt$WORK_DESCRIPTION, regex(single_family_words)),
                          MULTI = str_detect(permits_filt$WORK_DESCRIPTION, regex(multi_family_words)),
                          TH = str_detect(permits_filt$WORK_DESCRIPTION, regex(townhome_words)),
                          COACH = str_detect(permits_filt$WORK_DESCRIPTION, regex("COACH")))

write_csv(permits_filt, here("clean_data", "permits_filt.csv"))
```

########################
## Rent Data ##
########################
```{r fhfa}
fhfa <- read_csv(here("raw_data", "HPI_AT_BDL_tract.csv"), col_names = T)
fhfa <- right_join(fhfa, tract_data, by = c("tract" ="geoid10"))
fhfa <- mutate(fhfa, annual_change = as.numeric(annual_change),
                     hpi = as.numeric(hpi),
                     hpi1990 = as.numeric(hpi1990),
                     hpi2000 = as.numeric(hpi2000))
write_csv(fhfa, here("clean_data", "hpi.csv"))
```

```{r zillow}
zillow_shp <- readOGR(here("raw_data", "ZillowNeighborhoods-IL"), layer = "ZillowNeighborhoods-IL")
zillow_shp <- zillow_shp[which(zillow_shp$City == "Chicago"),]

zhvi_all <- read_csv(here("raw_data", "Neighborhood", "Neighborhood", "Neighborhood_Zhvi_AllHomes.csv"), col_names = T) %>%
  filter(RegionID %in% zillow_shp$RegionID) %>%
  gather(key = "Date",  value = "HVI", -c("RegionID", "RegionName", "City", "State", "Metro", "CountyName", "SizeRank")) %>%
  select(c("RegionID", "RegionName", "SizeRank", "Date", "HVI")) %>%
  mutate(Date = paste0(Date, "-15") %>% as.Date())

write_csv(zhvi_all, here("clean_data", "zhvi.csv"))
```


